---
format: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rgemmacpp

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of rgemmacpp is to provide a very basic wrapper around the CLI of Google's [gemma.cpp](https://github.com/google/gemma.cpp). 

The gemma.cpp code is included as a git submodule of this project and the existing cmake build system is used to compile all gemma.cpp related code as static libraries.

This is still very much a work in progress / proof of concept.



## Installation

You can install the development version of rgemmacpp from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("rundel/rgemmacpp")
```

In order to use the package you will need to download model weights and tokenizer - both of these are available from [the Gemma model page on Kaggle](https://www.kaggle.com/models/google/gemma) under `Model Variations |> Gemma C++`. See the gemma repo's [Quick Start - Step 1](https://github.com/google/gemma.cpp?tab=readme-ov-file#step-1-obtain-model-weights-and-tokenizer-from-kaggle) for more details.

Both the `2b` and `7b` models have been tested with this package - either the `2b-it-sfp` or `7b-it-sfp` model weights are suggested as a good place to start.


## Example

The following examples demonstrate the basics of prompting the 2b and 7b gemma model:

```{r}
library(rgemmacpp)
```

### 2b model

```{r}
m = gemma(
  tokenizer="~/Scratch/gemma/tokenizer.spm",
  compressed_weights="~/Scratch/gemma/2b-it-sfp.sbs",
  model="2b-it",
  multiturn="1"
)

r = m$prompt("What are top 5 places I should visit in Durham, NC?")

r = m$prompt("Which of the previous locations are best for kids?")
```

### 7b model

```{r}
m = gemma(
  tokenizer="~/Scratch/gemma/tokenizer.spm",
  compressed_weights="~/Scratch/gemma/7b-it-sfp.sbs",
  model="7b-it",
  multiturn="1"
)

r = m$prompt("What are top 5 places I should visit in Durham, NC?")

r = m$prompt("Which of the previous locations are best for kids?")
```
